Estimating the Initial Susceptible Fraction in Dengue Dynamics

### 2014-10-08 draft_lm1.tex
minor changes
added Poisson-binomial idea

To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
Carlo techiniques implemented in the Bayesian inference with Python 
(BIP)~\cite{pone2011} available at 
\url{http://code.google.com/p/bayesian-inference/}.
BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
scheme that efficiently samples from high-dimensional joint distributions using 
multiple adaptive chains running in parallel with delayed rejection.
In this study we used one chain per parameter, i.e, 3 chains for each run.
The chains were run until $5000$ samples were obtained after discarding $500$ 
burn-in samples.
Convergence of the parallel chains was verified at every 100 iterations by the 
calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
% Results and Discussion can be combined.
\section*{Results}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.

>>>>

To approximate the posterior in~(\ref{eq:S0post}) we use Markov chain Monte 
Carlo techiniques implemented in the Bayesian inference with Python 
(BIP)~\cite{pone2011} available at 
\url{http://code.google.com/p/bayesian-inference/}.
BIP uses a Differential Evolution Adaptive Metropolis (DREAM)~\cite{vrugt2008} 
scheme that efficiently samples from high-dimensional joint distributions using 
multiple adaptive chains running in parallel with delayed rejection.
In this study we used one chain per parameter, i.e, 3 chains for each run.
The chains were run until $5000$ samples were obtained after discarding $500$ 
burn-in samples.
Convergence of the parallel chains was verified at every 100 iterations by the 
calculation of the Gelman-Rubins' R convergence diagnostic~\cite{brooks1998}.
% Results and Discussion can be combined.
\section*{Results}
% We only support three levels of headings, please do not create a heading level below \subsubsection.
%\subsection*{Subsection 1}
\subsection*{Serotype proportion}
The susceptible population estimated in this paper is defined as individuals 
susceptible to at least one of the circulating serotypes in any given year. 
Let this Total number of susceptibles be denoted by $S_T$.

Let the number of susceptibles to a given dengue virus ($DENV_i$), $S_i$ be 
represented by a Binomial random variables $S_i \sim Bin(n, p_i)$ where $n$ is 
population size and p is the true proportion of susceptibles to $DENV_i$ in 
the population, with $i \in  \{1, 2, 3, 4\}$. In the absence of 
cross-immunity, i.e., the exposure to serotypes are independent, the 
total number of susceptibles $S_T=\sum_i S_i$  should follow a Poisson-Binomial 
distribution, $PoisBin(p_i,n)$, where
\begin{equation}
 P(S_T=k) = \sum \prod_i p_i \prod_i  (1-p_i)
\end{equation}

To elicit priors for the $p_i \sim Be(a,b)$, we used the proportion of 
serotyped cases at the end of each epidemic as the expected value for the $p_i$.

Let $S_{0T}$ and $S_{fT}$, be the total number of susceptibles before and 
after an epidemic, respectively

% the suceptibility can be manifested, in individuals, as one of 
% the following forms:
% \begin{equation}
%  S = \{S_{1}, S_{2}, S_{3}, 
% S_{4},S_{12},S_{123},S_{1234},S_{13},S_{14},S_{23},S_{234},S_{24},S_{34}\}
% \end{equation}

\section*{Discussion}
% Do NOT remove this, even if you are not including acknowledgments.
\section*{Acknowledgments}
LMC is grateful to Dr. Leonardo Bastos for useful discussions on the 
posterior inference for $R(t)$.
